---
title: 2015 Ferrybox QC log 
author: Tom Hull + Eleanor Haigh
editor_options: 
  chunk_output_type: console
---
# Set up
These arethe packages you will need to load into you library - each time you open your r project
```{r setup}
require(cefasMOS, quietly=T)
require(leaflet, quietly=T)
require(dygraphs, quietly=T)
require(ncdf4, quietly=T)
require(dplyr, quietly=T)
data("telids") # load list of instruments and telids from CefasMOS
Sys.setenv(tz="UTC") #set environement time zone to utc
source("C:\\Users\\EH08\\Documents\\June 19\\juneferrrybox\\checking scripts\\ferrybox_functions.R") #need to read in functions required for QC
```

# 10minute files to rdata PROCESS
You may have already done this step, see SOP xxxx
2016 data as an exmaple. 

```{r process_data, eval = F}
#fb2018 = read.ferrybox.10min("C:/FB/2015/", recursive = T, print_file = T)
#save(fb2015, file = "C:/FB/fb2015.rdata")
```

# Known issues
Below, insert the logged problems with the ferrybox over the year - taken from the FeryBox log txt file on the FB computer (can be accessed remotely).


2015 log; 

01-11-15 - FIXED new ship pump meant FB wasn't getting enough flow. Once tested empy valvues and flushed with fresh water, In and Main valves were ajusted and FB switched back o - BS
30-10-15 - Ferrybox not working after pump change - not sure why. Half full error message. Restarted PC, now says it has no GPS feed - BS
29-10-15 - PROBLEM Ferrybox ships main pump failed, engineers to replace - BS
24-10-15 - software had crashed and was stuck in Standby. Restarted - BS
03-10-15 - found fb computer switched off - when started jumpy mouse and no GPS signal - problem caused by computer recognising mouse as GPS. P. Bouch solved the problem unplugging GPS, rebooting computer and reconnecting GPS. E Capuzzo
09-09-15 - Primary GPS input changed to Seapathm heading now available. GPS ini file changed to $IN from $GP to match NEMA - TH
07-08-15 - "jumpy" mouse is caused by booting fb pc without sampler pc, sampler pc needs to boot first? - TH
07-08-15 - optode A and B changed to salinity setting 34 - TH
06-08-15 - Emergency stop button fitted, PC power moved to port CS4/14 - TH
06-08-15 - cytobuoy sampling tube fitted. - TH
05-08-15 - modifications made in software and plumbing for seafet, system does now not drain before wash cycle - TH
03-06-15 - Seapoint replaced (2757 removed and replaced with 3616) - TH
28-05-15 - POSSIBLE PROBLEM - Seapoint Flu sensitivity - TH
07-05-15 - PROBLEM - tsg pump turned off but ferrybox pump kept running = empty-auto-off not working - TH
05-05-15 - Phytoflash stuck 0 values, restarted module not work, complete FB reset did - TH
22-02-15 - seafet reinstalled with salt solution valves, pCO2 nmea module updated TH


# Pre Processing
Here the initial steps are taken; 
- Data loaded, duplicates removed.
- Telemetry id's are translated and Meisberg Ph data removed.
- Hull PRT variable is renamed SST for clarity.
- Missing data removed.

```{r pre_processing}
load("C:\\Users\\EH08\\Documents\\June 19\\juneferrrybox\\checking scripts\\2015-copy\\fb2015.rdata") #load ferrybox data into r
fb = copy(fb2015) # you will be working on a copy of the raw data
rm(fb2015) #remove original copy

  #de_duplication
anyDuplicated(fb)# returns >0 if any duplicates
fb = unique(fb) # removes duplicates (i.e where every row matches)

  # (optional) translate flags, this will need to be repeated later
fb[, Quality := as.numeric(Quality)]
  # translate telid
fb[, telid := as.numeric(telid)]
fb = merge(fb, telids, by = "telid")  #merge to add column describing instrument
  # strip out header details - these will be updated later
fb[, Cruise := NULL]
fb[, SIC := NULL]

  # remove variables we don't care about
  # don't worry about the pH sensor or sound velocity
fb = fb[!(telid %in% c(2000, 2001, 1193))]

 #unique(fb[,.(telid, variable, sensor_parameter)])

  # rename hull PRT if not already - Sea Surface Temperature 
fb[telid == 2064, variable := "SST"]

  # remove missing data
fb = fb[Count > 0]

  # check for badly calculated means?
nrow(fb[Mean > Max & !grepl("Vaisala", sensor_parameter)]) # known issues with Vaisala
nrow(fb[Mean < Min & !grepl("Vaisala", sensor_parameter)])

```

Complete the above step before continuing

# GPS

- work with track file, unique points rather than every row
- Check for impossible distances (dv/dt > 20knts)
- Stuck values
- add note of progess sheet if large amount of data removed ?add gps from conlog to recover

There can be times when the gps on the FB fails.
However in some circumstances the conlog data can be used.
The conlog has a PC system time, but also the GGL time

Next extract just the gps track.
- Check the speed and course values are reasonable.
- stuck values are checked for that lat, long and speed are not exactly the same
- if they are it means we're on DP and we don't want the data anyway


```{r gps}
  # plot cruise track and identify bad points and misslabeled cruises
#create object of just track information
track = fb[,.(Latitude = unique(Latitude), Longitude = unique(Longitude),
              Speed = unique(Speed), Course =  unique(Course)),
           by = list(dateTime)]  

anyDuplicated(track$dateTime) # check that GPS, Speed and Course are unique

hist(track$Speed) # check speed is sensible
hist(track$Course) # check course is sensible

  # check that speed does not exactly equal course - known problem
track[Speed == 0 & Course == 0, c("Speed", "Course") := NA]
track[Speed == Course, c("Speed", "Course") := NA]

ggplot(track)+geom_point(aes(dateTime,Course)) #plot to see if course looks ok
ggplot(track)+geom_point(aes(dateTime,Speed)) #plot to see if speed looks ok
#track[,Course:=NA] #remove course if needed 
#track[,Speed:=NA] #remove speed if needed

  # remove non-minute data (where timestamps are incorrect, not exactly on th minute)
track = track[lubridate::second(dateTime) == 0]

track[Longitude == 0 & Latitude == 0, c("Latitude", "Longitude") := NA] # remove 0,0 position if present

  # find other stuck values
track[Latitude == shift(Latitude) &
        Longitude == shift(Longitude) &
        Speed == shift(Speed),
      stuck := T]

track = track[is.na(stuck)] # removing stuck values
```

Check that no points are over land.
Eval=F is used as this is a computationally expensive task.

```{r over_land, eval=F}
#find any nas in coordinates- raise as issue if there are long sections

perc=(sum(is.na((track[,Latitude])))/length((track[,Latitude])))*100 #percentage of lat coordinates missing, if this number is large may need to explore conlog

#create version of track without nas in long and lat
test=na.omit(track,c("Latitude","Longitude"), invert=F)

#create spatial object using track file
points_sp = sp::SpatialPoints(test[,.(Longitude, Latitude)],
                           proj4string=sp::CRS("+proj=longlat +datum=WGS84"))
#warning re resolution can be ignored, using low resolution fine for visual check
#create spatial object of world map
poly = sp::spTransform(rworldmap::getMap("high"), sp::CRS("+proj=longlat +datum=WGS84"))

#compare spatial object to see where track falls onto land
overlaps = is.na(sp::over(points_sp, poly)$NAME)
#create new column that returns TRUE where overalp occurs
test[!overlaps, overland := T]

  # check where these points are that fall over land
if(any(test$overland)){
  leaflet(test[overland == T]) %>%
    addTiles() %>%
    addCircles(~Longitude, ~Latitude, popup = ~paste(dateTime))
}

#merge overland points back to original track object 
track=merge(track,test[,.(dateTime, overland)], by="dateTime", all.x=TRUE ) 

#then remove flagged overland points if needed
#track = track[is.na(overland)]
```

Checking and recalculating speed

```{r recalculate_speed}


  # recalculate speed and course from GPS
track = track[order(dateTime)] #order by date
track[, CourseC := ferrybox.course(dateTime, Latitude, Longitude, threshold = 61)]  #function calculates course from gps
track[, SpeedC := ferrybox.speed(dateTime, Latitude, Longitude, threshold = 61)] #function calculates speed from gps

track[SpeedC == 0]  # check no exactly zero speed

  # do calculated values match those the ferrybox has recorded, should have 1:1 relationship?
plot(track$Speed, track$SpeedC) #plot calculated speed vs ferrybox speed
plot(track$Course, track$CourseC) #plot ferrybox course vs actual course

nrow(track[is.na(Speed)]) #how many rows don't have speed
nrow(track[is.na(Course)]) #how many rows don't have course

#if there are lots of rows without speed or course, and calculated values agree well with those from the ferrybox then we can replace missing values with calculated
#track[is.na(Speed), Speed := SpeedC]
#track[is.na(Course),Course:=CourseC]

```

# Cruise names

filling in the cruise names throughout the year

```{r gap_minder}
#check for missing cruises in fb data 
  # gap minder, check for missing cruises
track = track[order(dateTime)] #order by date
track[, time_diff := as.numeric(dateTime) - shift(as.numeric(dateTime))]  #find time difference between current point in track and previous
track[time_diff > 60*60*24]  #find when the difference between one track point and the next is greater than a day, this should line up with dates of cruises, but there may be occasions where the ferrybox failed so large chunks are missing
```

```{r cruises}
cruises = fread(file = "C:/Users/EH08/Documents/June 19/juneferrrybox/checking scripts/2015-copy/cruises2015.csv")  # read in a csv of cruise dates for given year, must follow same format as previous years csvs 
cruises[, Start := as.POSIXct(Start, tz = "UTC", format = "%d/%m/%Y")] # format is the date format of your read in csv
#where cruises turnaround on the same day, time will need to be added to both the csv and this format section to avoid the same day being assigned two cruises
cruises[, End := as.POSIXct(End, tz = "UTC", format = "%d/%m/%Y")]

  # add cruise information to track
track[, dt2 := dateTime]
setkey(track, dateTime, dt2) # dt2 needed for foverlaps, fast matching in range
track = foverlaps(cruises, track, by.x = c("Start", "End")) #find were cruise dates overlap ferrybox data 
track = track[,.(dateTime, Latitude, Longitude, Course, Speed, Cruise)] #enure track contains only columns needed
track = track[!is.na(dateTime)] #remove rows in track that do not have a date and time associated with them 

  # allocate colors for leaflet
cruise_pal = colorFactor("Set1", unique(cruises$Cruise))

  # increase month (change 1:12, to individual intergers) to step though data-check that cruise tracks make sense 
leaflet(track[month(dateTime) == 1:12]) %>%
  addTiles() %>%
  addCircles(color = ~cruise_pal(Cruise), popup = ~paste(dateTime)) %>%
  addLegend(pal = cruise_pal, values = ~Cruise)

  # may nedd to edit cruises csv and repeat process as needed
```

# Flow lag

There is a time a lag between the water entering the boat inlet to actually entering the FB system, therefore we need to correct for this lag.
Typically 3 minutes, confirm lag is consistent over time between SST and seabird temperature.

```{r flow_lag}
dygraph(dcast.data.table(fb[telid %in% c(2064, 1190) & Quality == 0],
                         dateTime ~ variable, value.var = "Mean")) # using this dygraph you can visualise the flow lag by comapring hull prt to in ferrybo temperature

fb[in_fb_loop == T, dateTime := dateTime - 60*3] #apply lag of 3 minutes to data that is within the ferrybox loop

```


# combine track file

writing back the changes made in the track file to the fb data
```{r merge_track}
  # write back track and cruise data
fb[,c("Latitude", "Longitude", "Course", "Speed", "Cruise") := NULL] #remove columns from fb that are going to be replaced 
fb = merge(fb, track[,.(dateTime, Latitude, Longitude, Course, Speed, Cruise)],
      by = c("dateTime"), all.x = T) #merge track and fb objects
fb = fb[!is.na(Latitude) | !is.na(Longitude)] # remove where no position
rm(poly,points_sp, overlaps, cruise_pal) #remove objects no longer needed 
```

## Flow rates

For this step we check the flow rates and flwo pressure within the fb system is OK.

```{r flow_rates}

#flipping the datatable with specific varaible columns; FLOWIN, FWMAIN, FBPRES
flow = dcast.data.table(fb[variable %in% c("FLOWIN", "FWMAIN", "FBPRES")],
                        dateTime + Longitude + Latitude + Cruise ~ variable, value.var = "Mean") 
flow[, FRATIO := FLOWIN / FWMAIN] #calculate flow ratio for qc

#visually checking flow and pressure parameters
hist(flow$FBPRES)
hist(flow$FLOWIN, breaks = 60)
hist(flow[FRATIO > 0]$FRATIO, breaks = 60)

#assigning flow ok flags 
flow[, flow_ok := F] #here you create a new column where all values == F
flow[FRATIO > 2 & FRATIO < 5, flow_ok := T] # where the fratio is between 2 and 5, the flow is good
flow[FLOWIN < 15 | FBPRES < 150, flow_ok := F] # where flowin is less than 15 and pressure less than 150, flow is bad
summary(flow) #see how many rows are flagged as having bad flow 

#visual check
ggplot(flow[flow_ok == T]) +
  geom_point(aes(dateTime, FRATIO))

  # if we're happy write back  to main data file fb
fb = merge(fb, flow[,.(dateTime, Latitude,Longitude, Cruise, flow_ok)], by = c("dateTime", "Cruise", "Longitude", "Latitude"), all.x = T)
anyDuplicated(fb) #should return zero
nrow(fb[in_fb_loop == T & Quality == 0 & flow_ok == F]) #number of rows where flow was not ok
rm(flow) # tidy up the workspace environment

fb[flow_ok != T & in_fb_loop == T & Quality == 0, Quality := 6] # if in loop and flow bad add flag
fb[is.na(flow_ok) & in_fb_loop == T & Quality == 0, Quality := 6] # if in loop and no flow data, add flag
fb[, c("flow_ok") := NULL] #removing flow_ok column

```

# Wash cycle flagging

The ferrybox applies a flag (#64) to in-loop variables during the wash cycle.
It also applies a flag for a cool-down period after the wash, while the pipework cools.
This period is controlled in the ferrybox software.
Lets check the period is long enough and apply more flags if needed.

```{r post_wash}
  # just work with a single variable for clarity(seabird temp)
wash = fb[telid == 1190]
wash[grepl(ferrybox.errorcode(64),ferrybox.errorcode(Quality)),flag:=T] #if error code is 64 the value was recorded during cleaning so needs a wash flag
nrow(wash[flag==T]) #how many rows have a wash flag, may be zero as some is automatically removed 
wash[, time_diff := as.numeric(dateTime) - shift(as.numeric(dateTime))] #calculate difference between time points
wash[,temp_diff := Mean - (shift(Mean, n=1))] #calculate difference in temperature between points

nrow(wash[time_diff > 5*60]) #where time gap is greater than 5 mins-may indicate wash occured
nrow(wash[(abs(temp_diff)) > 1])  #where temp difference is greater than 1 between time steps
nrow(wash[time_diff >5*60 & (abs(temp_diff)) > 0.5]) #number of rows where time_diff is >5mins and temperature change +/- 0.5, these may indictae a rise in temperature due to a wash 


 # using 5 minute gap in data as indicator of wash, bit of a hack
wash[time_diff > 5*60, flag := F]
wash_times = wash[time_diff > 5*60]$dateTime
table(wash$flag)

wash[, id := month(dateTime)]
ggplot(wash[month(dateTime) == 1:12]) +
  geom_point(aes(dateTime, Mean, color=flag)) +
  scale_x_datetime(date_minor_breaks="day") +
  facet_wrap("id", scales="free")

ggplot(wash[month(dateTime) == 2]) +
  geom_point(aes(dateTime, Mean, color=flag)) +
  scale_x_datetime(date_minor_breaks="day")

```

If it is decided more data needs to be flagged after the acod wash, then follow the step below. 

Example; 2016 fb data
Looking at Jan and Feb it looks like post wash needs more flagging.

```{r flag_more_wash}

length(unique(as.Date(wash$dateTime)))
length(wash_times) # there should be roughly 1 wash per day, plus extras for going into port / servicing

wash[, wash_flag := lagged_flag(dateTime, wash_times, 600)] # add flag 10 mins after wash times as identified by a gap of 5 minutes or more
wash = wash[wash_flag == T]$dateTime #times where wash is falgged as happening 

  # write these flags back to the data
fb[in_fb_loop == T & dateTime %in% wash, Quality := Quality + 64] #match times where wash was happening to times in ferrybox data set and add quality flag 

ggplot(fb[variable == "TEMP" & month(dateTime) == 1]) +
  geom_point(aes(dateTime, Mean, color=as.factor(Quality))) #cycle through month by month to see if flagging is sufficient 

```

# Oxygen

There are various known issues with oxygen.This will stay at QA1.

```{r oxygen}

#use these plots to check for obviously unrealistic values, eg -ves
hist(fb[Quality == 0 & variable == "O2SAT"]$Mean)
range(fb[Quality == 0 & variable == "O2SAT"]$Mean)

#plot with time 
ggplot(fb[variable == "O2SAT" & Quality == 0]) +
  geom_line(aes(dateTime, Mean, color = Cruise))
#if any values seem unrealistic, remove


#test for stuck values and flag
#apply function to oxygen optode, ie telid 1097, to detect repeats of 3 or more where instrument was stuck
stuck_value(fb, 1097, 3)
fb=w #replace fb with table ouput from function #check product (w) is correct before replacing
rm(tester, y, w) #remove variables generated by function
#fb=fb[,telidspecreps:=NULL]
fb[stuck==T & telid==1097 & Quality==0, Quality:=Quality+1] #add qaulity flag where stuck values identified

```

#Quantum Yield 

Only basic level checks are made here, this will stay at QA1.
```{r Quantum yield}
#NOTE only basic level 1QA has been done here really. 
#use these plots and tests to identify obviously unrealistic values
hist((fb[variable == "QYIELD"]$Variance), breaks=1000)
hist(sqrt(fb[variable == "QYIELD"]$Variance), breaks=1000)

range(fb[Quality == 0 & variable == "QYIELD"]$Mean)

ggplot(fb[variable == "QYIELD"]) +
  geom_point(aes(dateTime, Mean, color=as.factor(Quality)))

ggplot(fb[variable == "QYIELD" & Quality==0]) +
  geom_point(aes(dateTime, Mean))

#test for stuck values and flag
stuck_value(fb, 1164, 3) #tests telid 1164 for repeats of 3 or more 
fb=w #replace fb with table output from function
rm(tester, y, w) #remove variables generated by function
fb=fb[,telidspecreps:=NULL] #remove variables generated by function
fb[stuck==T & telid==1164 & Quality==0, Quality:=Quality+1] #add qaulity flag where stuck values identified

```
  
# Met
Only basic level checks are made here, this will stay at QA1.

```{r windspd}

  # convert to SI
  # fb is actually in knots despite what it logs
fb[variable %in% c("WINDSP", "WINDRS"), Mean := Mean * 0.514]
fb[variable %in% c("WINDSP", "WINDRS"), Max := Max * 0.514]
fb[variable %in% c("WINDSP", "WINDRS"), Min := Min * 0.514]

#test for stuck values and flag #WINDSP
stuck_value(fb, 	2133, 3)  
fb=w
rm(tester, y, w)
fb=fb[,telidspecreps:=NULL]
fb[stuck==T & telid==	2133 & Quality==0, Quality:=Quality+1]

#test for stuck values and flag #WINDRS
stuck_value(fb, 	2134, 3)
fb=w
rm(tester, y, w)
fb=fb[,telidspecreps:=NULL]
fb[stuck==T & telid==	2134 & Quality==0, Quality:=Quality+1]


#use plots and checks to identify obviously unrealistic values, remove values if needed 
hist(fb[variable == "WINDRS" & Quality == 0]$Mean)

dygraph(fb[variable %in% c("WINDSP") & Quality==0 ,.(dateTime, Mean)])
dygraph(fb[variable %in% c("WINDRS") & Quality==0 ,.(dateTime, Mean)])

```


#Saving QA1 Complete data

```{r, QA1 complete}

#Add QAlevel flag to current data table 
fb[, QAlevel := 1]

#Saving completed QA1 data
fb[,stuck:=NULL] #remove stuck column
fb=as.data.table(fb) #make sure file is saved as data table 
save(fb, file="C:/Users/EH08/Documents/June 19/juneferrrybox/checking scripts/2015-copy/fb2015_QA1Complete.rdata")
#any other QA done must be worked from this saved rdata file 
```

#QA2 
# Temperatures
Now we start QA2 porcesses; checking and correcting parameters. Temperature, salinity and fluorescence are done first and are a higher priority. 

```{r temperature}
load("C:\\Users\\EH08\\Documents\\June 19\\juneferrrybox\\checking scripts\\2015-copy\\fb2015_QA1Complete.rdata")
anyDuplicated(fb) #check for duplicates, if there are any something has gone wrong in QA1

dygraph(dcast.data.table(fb[variable %in% c("SST", "TEMP") & (Quality == 0)])) #intial plot to look at data

fb[variable == "SST" & Mean <= 0 & Quality == 0, Quality := 1] # flag as below minimum
fb[variable == "TEMP" & Mean <= 0 & Quality == 0, Quality := 1] # flag as below minimum

  # any SST data above 25 degress is nonsense
fb[variable == "SST" & Mean > 25 & Quality == 0, Quality := 1] # flag as above maximum
fb[variable == "TEMP" & Mean > 25 & Quality == 0, Quality := 1] # flag

  # flag data where standard deviation too high
fb[variable %in% c("SST", "TEMP") & sqrt(Variance) > 0.2, Quality := 1024]

#check for stuck values and flag #SST
stuck_value(fb, 1190, 3)
fb=w
rm(tester, y, w)
fb=fb[,telidspecreps:=NULL]
fb[stuck==T & telid==1190 & Quality==0, Quality:=Quality+1]

#check for stuck values and flag #TEMP
stuck_value(fb, 2064, 3)  
fb=w
rm(tester, y, w)
fb=fb[,telidspecreps:=NULL]
fb[stuck==T & telid==2064 & Quality==0, Quality:=Quality+1]

#check for stuck values #optode TEMP
stuck_value(fb, 1098, 3) 
fb=w
rm(tester, y, w)
fb=fb[,telidspecreps:=NULL]
fb[stuck==T & telid==1098 & Quality==0, Quality:=Quality+1]
```

```{r temperature_SST_vs_TEMP}
#first make a new object called TEMP
  # self merge to match SST to TEMP
  # quality bits are split into SST and TEMP
TEMP = merge(fb[variable == "SST",
                .(dateTime, SST = Mean, SST.sd = sqrt(Variance),
                  PRT_SERIAL = serial, PRT_TELID=telid, Cruise, SST_Quality = Quality,Speed)],
             fb[variable == "TEMP",
                .(dateTime, TEMP = Mean, TEMP.sd = sqrt(Variance),
                  telid, serial, sensor_parameter,TEMP_Quality = Quality)],
             by = "dateTime", all = T) 

unique(TEMP[,.(sensor_parameter,serial,PRT_SERIAL)]) #check how many instruments used

  # plot all temperature parameters vs SST
  # temperatures should closely track SST -if they dont may need to flag out 
ggplot(TEMP[TEMP_Quality == 0 & SST_Quality == 0]) +
  geom_point(aes(TEMP, SST, color = paste(sensor_parameter, serial))) +
  coord_equal()

  # dTEMP check
  # rate of change of temperature should not be crazy for either
TEMP[, SST_dt := rate_of_change(dateTime, SST), by = serial] #calc rate of change
TEMP[, TEMP_dt := rate_of_change(dateTime, TEMP), by = serial]

hist(TEMP[grepl("Seabird", sensor_parameter) & SST_Quality == 0]$SST_dt) 
hist(TEMP[grepl("Seabird", sensor_parameter) & TEMP_Quality == 0]$TEMP_dt)

  # check rate of change again
ggplot(TEMP[grepl("Seabird", sensor_parameter) & SST_Quality == 0]) +
  geom_point(aes(dateTime, SST_dt))

  # apply flags for high rate of change
TEMP[SST_Quality == 0 & abs(SST_dt) > 0.020, SST_Quality := 1024]
TEMP[TEMP_Quality == 0 & abs(TEMP_dt) > 0.020, TEMP_Quality := 1024]

# flag out where prt temp is >0.5 deg higher than in ferrybox measuers of temperature-this indicated where prt was measuring out of air temp ie. remained retracted for times when ferrybox was on 
TEMP[SST > (TEMP+0.5), SST_Quality := 1024]


  # fit linear models between SST and TEMP # create a new object x to work with 
x = TEMP[TEMP_Quality == 0 & SST_Quality == 0 & !is.na(Cruise)]
plot(x$TEMP, x$SST) #should have 1:1 relationship
fit = lm(data = x, SST ~ TEMP)
#plot(fit)
coef(fit) #returns coefficients of fits 

fits = lapply(unique(x$Cruise),
              function(z) lm(SST ~ TEMP, data = x[Cruise == z,]
                             )) # apply lm to each cruise
names(fits) = unique(x$Cruise) 
sapply(fits, coef) #get coefficient of fit for each cruise
barplot(sapply(fits, function(x) sd(residuals(x))), main = "SD residuals by Cruise") #bar plot residuals of fits for each cruise 

#final visual check, onlt seabird temp has been qcd
ggplot(x[TEMP_Quality == 0 & SST_Quality == 0]) +
  geom_point(aes( SST,TEMP, color = paste(sensor_parameter, serial))) +
  facet_wrap("Cruise",scale="free")

```

Writing back TEMP flags to fb

```{r writeback_temp}
  # these should be the same length! This is just a check.
if(length(unique(TEMP$dateTime)) != length(unique(fb$dateTime))){
  warning("TEMP is the wrong size, something has gone wrong")
}

#write back qual flags with two seperate merges
#TEMP
fb=merge(fb, TEMP[,.(dateTime, TEMP_Quality, telid, serial)], by=c("dateTime", "telid", "serial"), all.x=T)
fb[!is.na(TEMP_Quality), Quality := Quality+TEMP_Quality]
fb[,"TEMP_Quality":=NULL]
#SST
fb=merge(fb, TEMP[,.(dateTime, SST_Quality, serial=PRT_SERIAL, telid=PRT_TELID)], by=c("dateTime", "telid", "serial"), all.x=T)
fb[!is.na(SST_Quality), Quality := Quality+SST_Quality]
fb[,"SST_Quality":=NULL]
fb=unique(fb)

#final visual check
dygraph(dcast.data.table(fb[telid %in% c(2064, 1190, 195, 1098) & Quality == 0],
                         dateTime ~ telid, value.var = "Mean"))

#remove leftover areas where prt out of water

```

```{r compare to profiler}

#set cruise id,parameter, min qa reached for profiler fetch function
cruiseID=NA
parameters = c("TEMP")
min_QA_reached = F
smartbuoy_db_name = "smartbuoydblive"
#set date range
after=min(fb$dateTime)
before=max(fb$dateTime)

#extract profiler data-takes a long time has already been run and saved-must be run again if new profilers added 

# ctd = profiler.fetch (cruiseID = cruiseID, after=after, before=before, parameters = parameters, db_name = smartbuoy_db_name, #min_QA_reached =    min_QA_reached)
    #ctd = ctd[depth > 3.5 & depth < 4.5] #only want prt depths
    #ctd[, `:=`(dateTime, as.POSIXct(round(as.numeric(dateTime)/60) * 
       # 60, origin = "1970-01-01", tz = "UTC"))] #round to nearest minute
    #ctd = ctd[, .(value = mean(value)), by = list(dateTime, site, 
        #latitude, longitude, sensor, par, cruise, profiler)] #only info of interest
    #ctd=ctd[grepl("CEND",cruise,ignore.case = TRUE)] #only want cend cruises
    #save(ctd,file="profilerdata.rdata")


setwd("C:/Users/EH08/Documents/June 19/juneferrrybox/checking scripts/2015-copy")
load("profilerdata.rdata")

    
    #get ferrybox sst/prt data needed
        proffb = fb[sensor_parameter == "ADAM PT100 PRT - parameter 1" & Quality==0] #get prt data
    proffb = proffb[, .(sensor_parameter, variable, dateTime, Latitude, Longitude, ferrybox_speed = Speed, 
        Mean)] #get only columns needed
    setkey(proffb, dateTime, variable)
    setkey(ctd, dateTime, par)
    
    #merge profiler and prt
        matched = merge(proffb, ctd, suffixes = c("_ferrybox", "_profiler"))
    setcolorder(matched, order(colnames(matched)))
    
    #get ferrybox temp/sebird data needed
    seabfb = fb[sensor_parameter== "Seabird SBE45 Thermosalinograph - parameter 1" & Quality==0] #get only ferrybox seabird temp
    seabfb = seabfb[, . (sensor_parameter,variable,dateTime,Latitude,Longitude,ferryox_speed=Speed, Mean)]
    setkey(seabfb,dateTime,variable)
    
    #mearge profiler and seabird
    matchedsb=merge(seabfb,ctd,suffixes=c("_ferrybox", "_profiler"))
    setcolorder(matchedsb,order(colnames(matchedsb)))
    
        #generate eqn for prt vs profiler
    eq = matched[, .(eq = lm_eqn(value, Mean), 
        xpos = min(value) + ((max(value - min(value))/2)), 
        ypos = max(Mean)), by = list(cruise, variable)]
    
    #generate equation for fb seabird vs profiler
        eq2 = matchedsb[, .(eq2 = lm_eqn(value, Mean), 
        xpos2 = min(value) + ((max(value - min(value))/2)), 
        ypos2 = max(Mean)), by = list(cruise, variable)]
    
    
    #plot prt vs both profiler and seabird
    plt = ggplot() + 
        geom_point(data=matched,aes(value, Mean)) + geom_point(data=matchedsb,aes(value,Mean)) + geom_smooth(data=matched, aes(value, Mean), method = "lm") + geom_smooth(data=matchedsb, aes(value, Mean), method = "lm",color="red") + geom_text(data = eq, 
        aes(x = xpos, y = ypos, label = eq), parse = T) + geom_text(data = eq2, 
        aes(x = xpos2, y = ypos2, label = eq2), parse = T)+ facet_wrap("cruise", scales = "free") + theme_bw() +labs( x = "Profiler FSI CT Module", y = "PRT (Blue) or Ferrybox seabird (Red)")
    print(plt)
    list(data = matched, plot = plt) #lists matched data between profiler, prt and seabird 
    
    #save ctd dta for later use
   # save(ctd,file="profilerdata.rdata")
    
```

```{r compare to rosette}
#find ctd files in network folder and list cruises that have data #change year in path
ctdfolders=list.files(path="Y:/SmartBuoyData/CTD - SBE/2017",full.names=TRUE)

   substrRight<- function(y, n){
  substr(y, nchar(y)-n+1, nchar(y))
   } #function to extract cruise name from directory
   
   cruises=lapply(ctdfolders,substrRight,n=10) # apply function to get list of cruises with ctd profiles
   cruises=unlist(cruises)# list->vector
   ctdpositions<-list()#predefine lists for read_ctd fucntion
   ctddata<-list()
   finalctddata<-list()

lapply(c(2:length(cruises)),read_ctd) #apply function to read in data for each cruise #cend_02 not qcd?

  stationdataf=rbindlist(finalctddata, fill=T) #list->data table 
  
      #get ferrybox sst/prt data needed
        proffb = fb[sensor_parameter == "ADAM PT100 PRT - parameter 1" & Quality==0] #where no quality flags
    proffb = proffb[, .(sensor_parameter, variable, dateTime, Latitude, Longitude, ferrybox_speed = Speed, 
        Mean, Cruise)]  #get columns needed
    setkey(proffb, dateTime, variable) 
    
    #get ferrybox temp/sebird data needed
    seabfb = fb[sensor_parameter== "Seabird SBE45 Thermosalinograph - parameter 1" & Quality==0]
    seabfb = seabfb[, . (sensor_parameter,variable,dateTime,Latitude,Longitude,ferryox_speed=Speed, Mean, Cruise)]
    setkey(seabfb,dateTime,variable)

    setkey(stationdataf, dateTime)# !!!!
    
        #merge ctd rosette and prt
        matched = merge(proffb, stationdataf, suffixes = c("_ferrybox", "_ctd"))
    setcolorder(matched, order(colnames(matched)))
    
        #mearge seabird and profiler
    matchedsb=merge(seabfb,stationdataf,suffixes=c("_ferrybox", "_ctd"))
    setcolorder(matchedsb,order(colnames(matchedsb)))

#plot
        #generate eqn for prt vs profiler
    eq = matched[, .(eq = lm_eqn(temperature, Mean), 
        xpos = min(temperature) + ((max(temperature - min(temperature))/2)), 
        ypos = max(Mean)), by = list(Cruise, variable)]
    
    #generate equation for fb seabird vs profiler
        eq2 = matchedsb[, .(eq2 = lm_eqn(temperature, Mean), 
        xpos2 = min(temperature) + ((max(temperature - min(temperature))/2)), 
        ypos2 = max(Mean)), by = list(Cruise, variable)]
    
    
    #plot prt vs both profiler and seabird
    plt = ggplot() + 
        geom_point(data=matched,aes(temperature, Mean)) + geom_point(data=matchedsb,aes(temperature,Mean)) + geom_smooth(data=matched, aes(temperature, Mean), method = "lm") + geom_smooth(data=matchedsb, aes(temperature, Mean), method = "lm",color="red") + geom_text(data = eq, 
        aes(x = xpos, y = ypos, label = eq), parse = T) + geom_text(data = eq2, 
        aes(x = xpos2, y = ypos2, label = eq2), parse = T) + facet_wrap("Cruise", scales = "free") +theme_bw() +labs( x = "CTD rosette", y = "PRT (Blue) or Ferrybox seabird (Red)")
    print(plt)
    list(data = matched, plot = plt)

```


# Salinity
Now we look at salinity. The  seabird salinometer is very stable, and usually has a relationship close to 1 with discrete sample. Best to check though. 

```{r salinity}
#plots to check values initially seem reasonable 
hist(fb[variable == "SAL" & Quality == 0]$Mean, breaks = 36)
hist(sqrt(fb[variable == "SAL" & Quality == 0]$Variance), breaks = 10)
range(fb[variable == "SAL" & Quality == 0]$Mean)

#visualise where low sals are-in mouth of river for eg, low salinity may be believable 
leaflet(fb[variable == "SAL" & Quality == 0 & Mean < 31]) %>%
  addTiles() %>%
  addProviderTiles(providers$Esri) %>% 
  addCircles(color = "red", popup = ~paste(dateTime ,Cruise ,Mean)) 

#check for stuck values and flag #SAL
stuck_value(fb, 1192, 3)  
fb=w
rm(tester, y, w)
fb=fb[,telidspecreps:=NULL]
fb[stuck==T & telid==1192 & Quality==0, Quality:=Quality+1]

```

```{r lims}
#discrete sample comparison
#use lims.fetch() function to extra salinity data in year range from the lims database
lims = lims.fetch(parameters=c("SAL"), after=min(fb$dateTime), before=max(fb$dateTime))#this is extracting data from lims after the first dateTime
lims=lims[depth>2 & depth<6] #only want depths near inlet
lims[, `:=`(dateTime, as.POSIXct(round(as.numeric(dateTime)/60) * 
        60, origin = "1970-01-01", tz = "UTC"))]   #round to nearest minute
lims=lims[,.(dateTime,latitude,longitude,gear, value,depth, survey,variable)] #only info of interest
lims=lims[grepl("CEND",survey,ignore.case = TRUE)] #only cend cruises
setkey(lims,dateTime,variable)


#You can extract data survey by surevy if needs be
#lims19_17SAL = lims[survey == "CEND 19/17"] # sal

#extract sal data from ferrybox
sal=fb[sensor_parameter=="Seabird SBE45 Thermosalinograph - parameter 3" & Quality==0] 
sal=sal[,.(sensor_parameter, dateTime, Longitude, Latitude, Speed, Mean, variable,serial)]
    setkey(sal,dateTime,variable)

#then merge fb and lims2
#from this new limsMerge object you can now run a regression of fb salinity against discrete salinity
xy= merge(sal, lims, suffixes=c("_ferrybox","_discrete"))

        #generate eqn for sal vs discrete for each serial no
    eq = xy[, .(eq = lm_eqn(value, Mean), 
        xpos = min(value) + ((max(value - min(value))/2)), 
        ypos = max(Mean)), by = list(serial, variable)]

    #plot discrete vs fb salinity
    ggplot(xy)+geom_point(aes(value,Mean))+geom_smooth(aes(value,Mean),method="lm") + geom_text(data = eq, 
        aes(x = xpos, y = ypos, label = eq), parse = T) + facet_wrap("serial", scales = "free") + theme_bw() +labs( x = "Discrete sample salinity (2-6m)", y = "Ferrybox seabird salinity")
    

#now apply factors and offsets to the data if needed
    coes=xy[,lm(Mean~value)] 
    #fb[sensor_parameter=="Seabird SBE45 Thermosalinograph - parameter 3", Mean:= (Mean-coes$coefficients[1])/coes$coefficients[2]]
    

```

```{r XXXX}
setkey(fb, NULL) # removing any keys we have set so far
anyDuplicated(fb)

```

# Flurometry

Next we check and correct fluorescence
There can be multiple fluoremeters used in one year. Will need to check what fluoroms have been used - see how many snsor ids have been recorded.

Example; 2016 
Two seapoint fluorometers were used during 2016: 2757 & 3616.
They have very difference responses.

```{r Flu}
unique(fb[grepl("FLU", variable), .(sensor_parameter,telid,variable,serial)]) #list number of fluorometes used

fb[variable %in% c("FLUORS", "FLUMAX") & Mean == 0 & Quality ==0, Quality := 1] # it is unlikely that mean will == 0 

  # you are comparing the flumax (a known constant fluoremeter) to the seabird fluoroms
#we are just checking the relationship/behaviour of the fluoroms here - not correcting anything yet.

#check for stuck values nad flag #FLUORS
stuck_value(fb, 144, 3)  
fb=w
rm(tester, y, w)
fb=fb[,telidspecreps:=NULL]
fb[stuck==T & telid==144 & Quality==0, Quality:=Quality+1]

#check for stuck values and flag #FLUMAX
stuck_value(fb, 1161, 3) 
fb=w
rm(tester, y, w)
fb=fb[,telidspecreps:=NULL]
fb[stuck==T & telid==1161 & Quality==0, Quality:=Quality+1]

#Make a new datatable called FLU 
FLU = dcast.data.table(fb[variable %in% c("FLUORS", "FLUMAX") & 
Quality == 0],dateTime + Cruise ~ variable + serial, value.var = c("Mean")) #fun agg as most are not identified as repeats

lm(data = FLU, FLUMAX_2500127 ~ FLUORS_2757) # you are just looking at the relationship here between flumax and fluors
lm(data = FLU, FLUMAX_2500127 ~ FLUORS_3616) # these coefs are not applied 


#now create daattabe x to work with, containing only good data from fluorometers
x = fb[variable %in% c("FLUORS", "FLUMAX") & Quality == 0]
x = dcast.data.table(x, dateTime ~ variable, value.var = "Mean", fun.aggregate=mean)
x = merge(x, fb[variable == "FLUORS",.(dateTime, serial, Cruise)], by = "dateTime")

x[, id := month(dateTime)] # id for each month 



#plotting fluors with flumax to see relationship
ggplot(x[FLUMAX < 10000 & id %in% c(1:12)], aes(FLUMAX, FLUORS)) +
  geom_point(aes(shape = serial)) +
  geom_point(aes(FLUMAX, FLUORS), color = "red", size=0.25) +
  geom_smooth(aes(color = serial), method = lm)

#visualise by month
ggplot(x[FLUORS < 50], aes(FLUMAX, FLUORS)) +
  geom_point(aes(color = serial), size = 0.025) +
  facet_wrap("id")

fits = lapply(1:12, function(z) lm(FLUORS ~ FLUMAX, data = x[id == z,])) # apply lm to each element(month) - to get an idea of behaviour over the year
sapply(fits,coef)
#the flumax is around x1000 larger than fluorometer values. We are just exploring the data here
lm(data=x[serial==2757], FLUMAX/1000 ~ FLUORS)
lm(data=x[serial==3616], FLUMAX/1000 ~ FLUORS)

dygraph(FLU[,.(dateTime, Scaled_FLUMAX = FLUMAX_2500127/1000, FLUORS_2757, FLUORS_3616)]) # this is a good visual tool

#if relationship between given fluorometer and phytoflash is consistent then can continue

  # check for high variance and flag if needed
fb[variable == "FLUORS" & Quality == 0 & sqrt(Variance) > 1, ]
```

After exploring the fluorescence data, we can now apply coefs from the appropraite bead calibrations - this will give a corrected fluorescence value

```{r flu_corrections}

#Applying standardision from bead calibrations as found in 'standardisations' spreadsheet
fb = fb[variable == "FLUORS" & serial == 2757, Mean := 0.0767 + Mean * 1.967] 
fb[variable == "FLUORS" & serial == 3616, Mean := -0.0483 + Mean * 0.983] 

fb[variable == "FLUORS" & serial == 2757, unit := "Standardised Arb.Units"] # change units of measurement 
fb[variable == "FLUORS" & serial == 3616, unit := "Standardised Arb.Units"] 

#Visual check of the corrected data over the year
ggplot(fb[variable == "FLUORS" & Quality==0], aes(dateTime, Mean)) +
  geom_point(aes(color = serial))
dygraph(fb[variable=="FLUORS" & Quality==0,.(dateTime, Mean)])

dygraph(dcast.data.table(fb[variable == "FLUORS" & Quality == 0], dateTime ~ serial, value.var = "Mean", fun.aggregate=mean))

```
  
# Backscatter

```{r obs}

#initial plots to check data is reasonable, remove data if not
ggplot(fb[variable == "FTU" & Quality == 0]) +
  geom_line(aes(dateTime, Mean, color = Cruise))
ggplot(fb[variable == "FTU" & Quality == 0]) +
  geom_histogram(aes(Mean), binwidth = 1) + scale_y_log10()

summary(fb[variable == "FTU" & Quality == 0,.(Mean, sd = sqrt(Variance))])

  # flag if high variance
fb[variable == "FTU" & Quality == 0 & sqrt(Variance) > 10, Quality := 1]

#check for stuck values and flag #FTU
stuck_value(fb, 16, 3) 
fb=w
rm(tester, y, w)
fb=fb[,telidspecreps:=NULL]
fb[stuck==T & telid==16 & Quality==0, Quality:=Quality+1]


```

# PAR

```{r par_variance}
#inital plots to explore data, remove suspiscious points if found
dygraph(fb[variable == "PAR" & Quality == 0, .(dateTime, Mean)])
hist(sqrt(fb[variable == "PAR"]$Variance), breaks=1000)
range(fb[Quality == 0 & variable == "PAR"]$Mean)
dygraph(dcast.data.table(fb[variable == "PAR" & Quality == 0], dateTime ~ Cruise, value.var = "Mean"))

fb[sqrt(Variance) > 100, Quality := 1024] #flag high variance 

#check for stuck values and flag #PAR
stuck_value(fb, 96, 3)  
fb=w 
rm(tester, y, w)
fb=fb[,telidspecreps:=NULL]
fb[stuck==T & telid==96 & Quality==0, Quality:=Quality+1]

```
  
#complete QA2

```{r, completed QA2}
#adding QAlevel column

fb[variable %in% c("FLUORS", "TEMP", "SST", "SAL", "FTU", "PAR"), QAlevel := 2] # Where the variable is equal to etc give the row a QAlevel of 2. Add any variables as appropriate

#write fB object of completed qa2a data
fb[,stuck:=NULL] #remove stuck column
fb=as.data.table(fb) #ensure data is saved as data table 
save(fb, file="C:\\Users\\EH08\\Documents\\June 19\\juneferrrybox\\checking scripts\\2015-copy\\2015QA2_complete.rdata")
load("C:\\Users\\EH08\\Documents\\June 19\\juneferrrybox\\checking scripts\\2015-copy\\2015QA2_complete.rdata")
```

#QA3

#Deriving SPM 
```{r derive_spm}
#using realtionship as derived from full ferrybox series prior to 2019
#load in coefficients from script that utilises whole ferrybox series and lims discrete
load("C:\\Users\\EH08\\Documents\\June 19\\juneferrrybox\\checking scripts\\Discrete calibrations (full series)\\spm_coefficients.Rdata")
turbs=fb[(telid==16)] #isolate turbidity data
turbs[,telid:=17] #assign new telid for new spm variable
turbs[,sensor_parameter:=telids[telid==17,sensor_parameter]] #assign sensor parameter name associated with telid 17
turbs[,variable:="SPM"] #name new variable 
turbs[,unit:="g m-3"] #add units
turbs[,Variance:=NA] #remove variance-not applicable for derived product
turbs[,QAlevel:=3] #update QA level
turbs[,Mean:= (Mean-coes$coefficients[1])/coes$coefficients[2]] #apply coefficients from ftu-spm regression
fb=rbind(fb,turbs) #add derived spm to main fb object
rm(turbs) #remove spm object
```

#Deriving Chl
**Chlorophyll not currently derived QA3**
  
```{r derive chlorophyll}
fb[telid==144,unique(serial)] #find unique fluorometer serial numbers

#discrete sample comparison
#use lims.fetch() function to get chlorphyll data in year range from the lims database
lims = lims.fetch(parameters=c("CHLOROPHYLL"), after=min(fb$dateTime), before=max(fb$dateTime))#this is extracting data from lims after the first dateTime
lims=lims[depth>3 & depth<5] #only want depths near inlet
lims[, `:=`(dateTime, as.POSIXct(round(as.numeric(dateTime)/60) * 
        60, origin = "1970-01-01", tz = "UTC"))]   #round to nearest minute
lims=lims[,.(dateTime,latitude,longitude,gear, value,depth, survey,variable)] #only info of interest
lims=lims[grepl("CEND",survey,ignore.case = TRUE)] #only cend cruises
setkey(lims,dateTime)

#extract fl data from ferrybox
fl=fb[telid==144 & Quality==0]
ggplot(fl)+
  geom_point(aes(dateTime,Mean))
fl=fl[,.(sensor_parameter, dateTime, Longitude, Latitude, Speed, Mean, variable,serial, Cruise)]
    setkey(fl,dateTime)
    
#then merge fb and lims
#from this new limsMerge object you can now run a regression of fb fluorescence against discrete chlorophyll
xy= merge(fl, lims, suffixes=c("_ferrybox","_discrete"))
    
#plots to examine covergae of chlorophyll values
ggplot(xy)+
  geom_point(aes(Mean,value,colour=survey))

ggplot(xy)+
  geom_point(aes(dateTime,Mean), colour="blue")+
  geom_point(aes(dateTime,value), colour="green")

    eq = xy[, .(eq = lm_eqn(Mean,value), 
        xpos = min(Mean) + ((max(Mean - min(Mean))/2)), 
        ypos = max(value))] #equation for all fluorescence vs chlorophyll data from the year

    ggplot(xy)+geom_point(aes(Mean,value))+geom_smooth(aes(Mean,value),method="lm") + geom_text(data = eq, 
        aes(x = xpos, y = ypos, label = eq), parse = T) + facet_wrap("serial", scales = "free") + theme_bw() +labs( x = "Ferrybox fluorescence", y = "Discrete sample chl")
    
        xy=xy[!is.na(Cruise)]
    #eqauation and plot by cruise
        eq = xy[, .(eq = lm_eqn(Mean,value), 
        xpos = min(Mean) + ((max(Mean - min(Mean))/2)), 
        ypos = max(value)), by = Cruise]

    #plot discrete vs fb by cruise
    ggplot(xy)+geom_point(aes(Mean,value))+geom_smooth(aes(Mean,value),method="lm") + geom_text(data = eq, 
        aes(x = xpos, y = ypos, label = eq), parse = T) + facet_wrap("Cruise", scales = "free") + theme_bw() +labs( x = "Ferrybox fluorescence", y = "Discrete sample chl")
    
    fits = lapply(unique(xy$Cruise),
              function(z) lm(value ~ Mean, data = xy[Cruise == z,]
                             )) # apply linear regression to each cruise
names(fits) = unique(xy$Cruise)

#looks like cend 3  8  18 and 22 have good relationship-so list below
cruises=c("CEND_03_15", "CEND_08_15", "CEND_18_15", "CEND_22_15")
#define function to extract coefficients from 'fits' and apply for given cruise
get_coeffs<-function(x){
  
  int=fits[[x]][["coefficients"]][["(Intercept)"]]
  slope=fits[[x]][["coefficients"]][["Mean"]]
  chls=fb[Cruise==x & telid==144 & Quality==0]
    chls=chls[,':=' (Mean=(Mean*slope)+int, variable="CHL", QAlevel=3, unit="ug/l", telid=145, sensor_parameter="Seapoint chlorophyll flurometer (second)")]
}

chl=list() #predefine list to store lapply results
chl=lapply(cruises, get_coeffs) #apply get_coeffs function to list of selected cruises
chls=rbindlist(chl) #concatenate results
fb=rbindlist(list(fb, chls)) #add results of discrete deriation to main fb object

plot(chls$dateTime, chls$Mean)

##this step is not complete-a generic relationship is being derived to be applied to other surveys
```


#complete QA3

```{r, completed QA3}
#adding QAlevel column

fb[variable == c("CHL", "SPM"), QAlevel := 3] # Where the variable is equal to etc give the row a QAlevel of 3. Add any variables as appropriate

#write fb object of completed QA3 data
#save(fb, file = "C:\\Users\\SH16\\OneDrive - CEFAS\\R\\FerryBox\\FB 2016\\Rdata\\QA2a\\fb_QA3_Complete.rdata")
```

# Create NetCDF

## Variables:
to publish: SST, SAL, FTU, FLUORS, SPM, CHL, PAR

```{r netCDF}
load("C:\\Users\\EH08\\Documents\\June 19\\juneferrrybox\\checking scripts\\2015-copy\\2015QA3_complete.rdata")
library(ncdf4)
library(oce)
fb = fb[QAlevel>=2 & Quality==0] #only qa 2 or higher 
anyDuplicated((fb))
serials=fb[,.(unique(serial))]
setnames(serials,"V1","serial")
instlist=unique(merge(serials, fb[,.(variable, telid,serial)], by="serial",all.y=F))

#get lats and longs for unique timestamps
timedat=fb[,.(dateTime, Longitude, Latitude)]
timedat=timedat[!is.na(dateTime),lapply(.SD,mean,na.rm=T), by= list(dateTime)]
cruisedat=fb[,.(dateTime,Cruise)]
cruisedat=unique(cruisedat, by=c("dateTime", "Cruise"))
datetimedat=merge(timedat, cruisedat, by="dateTime")

# get parameter info from csv
fb_params=as.data.table(read.csv("C:\\Users\\EH08\\Documents\\June 19\\juneferrrybox\\checking scripts\\ferrybox parameters.csv",header=T,stringsAsFactors = F))

#make netcdf
#define dimensions
#dimensions will be trajectory
traj_dim=ncdim_def("trajectory", "", 1, longname="Cefas Endeavour Ferrybox Data 2015")
obs_dim=ncdim_def("obs", "", 1:length(unique(as.numeric(fb$dateTime))) )

#define variables
#latitude
nc_lat=ncvar_def("latitude", "degrees_north", list(traj_dim, obs_dim), missval=-99999,prec='double')
#longitude
nc_lon=ncvar_def("longitude", "degrees_east", list(traj_dim, obs_dim), missval=-99999,prec='double')
#crs
nc_crs=ncvar_def("crs","",list())
#depth
nc_depth=ncvar_def("depth", "metres", list(traj_dim, obs_dim), missval=-99999, prec='double')
#time
nc_time=ncvar_def("time", "seconds since 1970-01-01 00:00:00 UTC", list(traj_dim, obs_dim), missval=-99999, prec='double')
#cruise  need special treatment as is a character string
char_dim=ncdim_def("nchar","",1:20, create_dimvar = F)
nc_cruise=ncvar_def("cruise", "", list(char_dim, obs_dim), prec='char')

test_list=(list(nc_time,nc_lat, nc_lon, nc_crs, nc_depth, nc_cruise))

#define other variables
for (var in fb_params$telid){
  print(var)
  if (any(unique(fb$telid==var))){
  metadata=fb_params[telid==var]
  nc_var=ncvar_def(metadata$parameter, metadata$units, list(traj_dim, obs_dim), missval=metadata$X_fillvalue, longname = metadata$long_name)#define varible for given var
  assign(paste("nc_", metadata$parameter,sep=""), nc_var) #rename nc_var to indicate which variable this object conscerns
  test_list[[length(test_list)+1]]<-eval(parse(text = (paste("nc_", metadata$parameter,sep="")))) #add to list which will later be used in netcdf creation 
rm(nc_var)
if (var!=145){
  if (var!=17){
nc_var=ncvar_def(metadata$instid, "", list(traj_dim, obs_dim), longname = metadata$sensor_name)
assign(paste("nc_", metadata$instid,sep=""), nc_var) 
test_list[[length(test_list)+1]]<-eval(parse(text = (paste("nc_", metadata$instid,sep=""))))
}
}
rm(nc_var)
}
}


new_nc=nc_create("Fb2015.nc", test_list)

#now add data to netcdf
#time
ncvar_put(new_nc, nc_time, unique(fb$dateTime))
#ncatt_put
ncatt_put(new_nc, nc_time, "standard_name", "time")
ncatt_put(new_nc, nc_time, "calendar", "gregorian")
ncatt_put(new_nc, nc_time, "axis", "T")

#get latitudes by unique datetime 
#latitude
ncvar_put(new_nc, nc_lat, datetimedat[,Latitude])
#ncatt_put
ncatt_put(new_nc, nc_lat, "standard_name", "latitude")
ncatt_put(new_nc, nc_lat, "units", "degrees_north")
ncatt_put(new_nc, nc_lat, "axis", "Y")
ncatt_put(new_nc, nc_lat, "valid_min", "-90")
ncatt_put(new_nc, nc_lat, "valid_max", "90")
ncatt_put(new_nc, nc_lat, "grid_mapping", "crs")
ncatt_put(new_nc, nc_lat, "axis", "Y")

#longitude
ncvar_put(new_nc, nc_lon, datetimedat[,Longitude])
ncatt_put(new_nc, nc_lon, "standard_name", "longitude")
ncatt_put(new_nc, nc_lon, "units", "degrees_east")
ncatt_put(new_nc, nc_lon, "axis", "X")
ncatt_put(new_nc, nc_lon, "valid_min", "-180")
ncatt_put(new_nc, nc_lon, "valid_max", "180")
ncatt_put(new_nc, nc_lon, "grid_mapping", "crs")
ncatt_put(new_nc, nc_lon, "axis", "X")

#crs
ncatt_put(new_nc, nc_crs, "grid_mapping_name", "latitude_longitude")
ncatt_put(new_nc, nc_crs, "espg_code", "ESPG:4326")
ncatt_put(new_nc, nc_crs, "longitude_of_prime_meridian", 0.0)
ncatt_put(new_nc, nc_crs, "longitude_of_prime_meridian", 6378137.0)
ncatt_put(new_nc, nc_crs, "inverse_flattening", 298.257223563)

#depth
ncvar_put(new_nc,nc_depth, rep(4, times=length(unique(fb$dateTime))))
ncatt_put(new_nc, nc_depth, "standard_name", "depth")
ncatt_put(new_nc, nc_depth, "units", "metres")
ncatt_put(new_nc, nc_depth, "long_name", "nominal depth of water inlet")
ncatt_put(new_nc, nc_depth, "axis", "Z")

#cruise
ncvar_put(new_nc,nc_cruise, datetimedat[,Cruise])
ncatt_put(new_nc, nc_cruise, "long_name", "survey on which sample was recorded")

# need to fill value where varibale does not exist for timestamp!!!

for (var in fb_params$telid){
  print(var)

  if (any(unique(fb$telid==var))){
  metadata=fb_params[telid==var]
  vardata=fb[telid==var]
  vardata=dcast.data.table(vardata, dateTime+serial~variable, value.var=c("Mean"), fun.aggregate = mean) #get data for given variable
  vardata=merge(vardata, datetimedat, by="dateTime", all.y=T) #merge with dtetime vector 
  vardata[is.na(vardata)]=-99999 #replace nas with -99999
  varserials=vardata[,2]
  vardata=vardata[,3]
  
  #add new var data to netcdf
  ncvar_put(new_nc, eval(parse(text = (paste("nc_", metadata$parameter,sep="")))), as.matrix(vardata))
    ncatt_put(new_nc, eval(parse(text = (paste("nc_", metadata$parameter,sep="")))), "standard_name", metadata$standard_name)
  ncatt_put(new_nc, eval(parse(text = (paste("nc_", metadata$parameter,sep="")))), "instrument", metadata$instid)
      ncatt_put(new_nc, eval(parse(text = (paste("nc_", metadata$parameter,sep="")))), "valid_min", metadata$valid_min)
    ncatt_put(new_nc, eval(parse(text = (paste("nc_", metadata$parameter,sep="")))), "valid_max", metadata$valid_max)
    ncatt_put(new_nc, eval(parse(text = (paste("nc_", metadata$parameter,sep="")))), "grid_mapping", "crs")
    
      #add varb instrument to netcdf-but not for chl or spm
  if (var!=145){
    if (var!= 17){
    ncatt_put(new_nc, eval(parse(text = (paste("nc_", metadata$instid,sep="")))), "nodc_name", metadata$nodc_name)
    ncatt_put(new_nc, eval(parse(text = (paste("nc_", metadata$instid,sep="")))), "make_model", metadata$sensor_make)
    ncvar_put(new_nc, eval(parse(text = (paste("nc_", metadata$instid,sep="")))), as.matrix(varserials))
    ncatt_put(new_nc, eval(parse(text = (paste("nc_", metadata$instid,sep="")))), "instrument precision", as.character(metadata$precision))
    }
  }
  }
}

yr=2015

#write global attributes-to develop-not working yet
ncatt_put(new_nc, 0, "id",paste(paste(yr,"fb",sep="")))
ncatt_put(new_nc, 0, "title", paste("Cefas Endeavour ferrybox time series from", paste(yr,collapse=", ")))
ncatt_put( new_nc, 0, "ncei_template_version","NCEI_NetCDF_Trajectory_Incomplete _Template_v2.0")
ncatt_put(new_nc, 0, "featureType","trajectory")
ncatt_put(new_nc, 0, "summary",paste("Ferrybox ", paste(sel_cruise,collapse=", "),". Parameters include: sea surface temperature, sea surface salinity, chlorophyll concentration, suspended particulate matter, photosynthetically active radiation"))
ncatt_put(new_nc, 0, "keywords","EARTH SCIENCE > Oceans > Ocean Temperature > Water Temperature, EARTH SCIENCE > Oceans > Ocean Optics > Photosynthetically Active Radiation, EARTH SCIENCE > Oceans > Ocean Optics > Turbidity, EARTH SCIENCE > Oceans > Ocean Optics > Fluorescence, EARTH SCIENCE > Oceans > Salinity/Density > Salinity, EARTH SCIENCE > Oceans > Ocean Pressure > Water Pressure, EARTH SCIENCE > Oceans > Ocean Chemistry > Oxygen, EARTH SCIENCE > Oceans > Ocean Chemistry > Chlorophyll, EARTH SCIENCE > Oceans > Ocean Chemistry > Suspended Solids")
ncatt_put(new_nc, 0, "keywords_vocabluary","GCMD:GCMD Keywords")
ncatt_put(new_nc, 0, "Conventions","CF-1.6, ACDD-1.3")
ncatt_put(new_nc, 0, "naming_authority","uk.co.cefas")
ncatt_put(new_nc, 0, "source","4H JENA Ferrybox system")
ncatt_put(new_nc, 0, "processing_level","Quality Controlled (Level 2")
ncatt_put(new_nc, 0, "licence","OGLv3.0 https://www.nationalarchives.gov.uk/doc/open-government-licence/version/3/") 
ncatt_put(new_nc, 0, "standard_name_volcabulary","CF Standard Name Table v39")
ncatt_put(new_nc, 0, "date_created", strftime(lubridate::now(tzone="UTC"), "%Y-%M-%dT%H:%M:%SZ"))
ncatt_put(new_nc, 0, "creator","Eleanor Haigh")
ncatt_put(new_nc, 0, "creator_email","eleanor.haigh@cefas.co.uk")
ncatt_put(new_nc, 0, "creator_url","http://www.cefas.co.uk")
ncatt_put(new_nc, 0, "institution","Centre for Environment Fisheries and Aquaculture Science")
ncatt_put(new_nc, 0, "acknowledgment","")
ncatt_put(new_nc, 0, "project","DP705")
ncatt_put(new_nc, 0, "geospatial_lat_min",min(datetimedat$Latitude[datetimedat$Latitude!=-99999]))
ncatt_put(new_nc, 0, "geospatial_lat_max",max(datetimedat$Latitude[datetimedat$Latitude!=-99999]))
ncatt_put(new_nc, 0, "geospatial_lon_min",min(datetimedat$Longitude[datetimedat$Longitude!=-99999]))
ncatt_put(new_nc, 0, "geospatial_lon_max",max(datetimedat$Longitude[datetimedat$Longitude!=-99999]))
ncatt_put(new_nc, 0, "geospatial_lat_units","degrees_north")
ncatt_put(new_nc, 0, "geospatial_lon_units","degrees_east")
ncatt_put(new_nc, 0, "time_coverage_start", as.character(min(datetimedat$dateTime[datetimedat$dateTime!=-99999])))
ncatt_put(new_nc, 0, "time_coverage_end",as.character(max(datetimedat$dateTime[datetimedat$dateTime!=-99999])))
ncatt_put(new_nc, 0, "cdm_data_type","Trajectory")
ncatt_put(new_nc, 0, "references","doi::")


nc_sync(new_nc)

nc_close(new_nc)
```



#Steps that require developement

All step below need furhter development and testing. None of the below have been applied to any FB data. 
This is due to time constraints, and lack in availbale data - for example conlog
These steps are beneficial but not fundamental
##Conlog step

Note that this Conlog step is not finalised - and has not been completed for any of the years yet. Therefore, there will be missing fb data (due to lack of gps). 

The aim is to implement this step at a later date.It will be in QA1 after GPS step. 

The conlog file is extracted fromt he Endeavour, wiher ask Bill Meadows for this file or ask the Master on the Endeavour.

```{r read_conlog, eval=F}
conlog_files = list.files("conlog/", pattern="*.csv", recursive=T, full.names=T)
conlog = lapply(conlog_files, read.ferrybox.conlog)
conlog = rbindlist(conlog)
conlog = unique(conlog)

  # reshape and calculate positions
conlog = dcast.data.table(conlog, Time ~ parameter, value.var="value", fun.aggregate=mean)

  # need to handle dates bring wrong around midnight
conlog[, decimal_time := (hour(Time)*3600 + minute(Time)*60 + second(Time)) / 86400]
conlog[, GLL_date := as.Date(Time)]
conlog[(decimal_time - GLL_time) > 0.9, GLL_date := GLL_date + 1] # correct for midnight
conlog[, fix_time := GLL_time*86400 + as.numeric(as.POSIXct(GLL_date, tz="UTC"))]
conlog[, fix_time := as.POSIXct(fix_time, origin="1970-01-01", tz="UTC")]

  # conlog is instant value on the minute
conlog[, dateTime := as.POSIXct(ceiling(as.numeric(fix_time)/60)*60, origin="1970-01-01", tz="UTC")]

conlog[, Lat := convert_latlong(LAT_deg, LAT_min)]
conlog[LON_hem_as_num == 90, EW := "E"]
conlog[LON_hem_as_num == 270, EW := "W"]
conlog[, Lon := convert_latlong(LON_deg, LON_min, EW)]
conlog = conlog[, lapply(.SD, median), .SDcols=c("Lat", "Lon"), by=dateTime]
save(conlog, file="conlog_2016.rdata")
```

```{r conlog_verification}
  # check processing is correct
load("conlog_2016.rdata")

conlog = unique(conlog) # remove duplicates
  # check it's sane
conlog = conlog[Lat %between% c(40, 65) & Lon %between% c(-30, 30)]
conlog[Lon == 0 & Lat == 0, c("Lat", "Lon") := NA] # remove 0 position
# odd bug in conlog data for 2016, time at 10:46 is wrong
conlog[hour(dateTime) == 10 & minute(dateTime) == 46, c("Lat", "Lon") := NA]

conlog[, c("dLat", "dLon") := list(sd(Lat), sd(Lon)), by=round_minute(dateTime, 10)]
  # remove stuck conlog values
conlog[dLat < 1e-05 & dLon < 1e-05, c("Lat", "Lon") := NA]
conlog = na.omit(conlog)

  # check conlog isn't doing stupid things around 0 lat
ggplot(conlog[Lon %between% c(-0.05, 0.05)]) +
  geom_point(aes(Lon, Lat, color="conlog"))

  # merge the track and conlog so we can compare timestamps / positions
x = merge(track, conlog, by="dateTime")[, .(dateTime, Latitude, Longitude, Speed, Lat, Lon, stuck)]

  # distance between conlog/ferrybox 
x[, dist_conlog := geosphere::distGeo(cbind(Longitude, Latitude), cbind(Lon, Lat))]
max(x$dist_conlog, na.rm=T)

  # maximum expected disparity ~463m (0.514*60*15) 15 knots to meters per minute
x[dist_conlog > 500] # this is indicative of clock drift

# ggplot(x) + geom_point(aes(dateTime, dist_conlog))
ggplot(x) + geom_point(aes(dateTime, dist_conlog))

lm(x$Lat ~ x$Latitude)
```


```{r fill_from_conlog}
track = merge(track, conlog[,.(dateTime, Lat, Lon)], by="dateTime", all=T)
track[is.na(Latitude), Latitude := Lat]
track[is.na(Longitude), Longitude := Lon]
track = track[!is.na(Longitude) | !is.na(Latitude),.(dateTime, Latitude, Longitude, Speed, Course)]

  # check for unique position at each timestamp
track[, n := 1:nrow(.SD), by=dateTime]
track[n > 1] # yay

```

# Stuck values
This requires more clarification. 

```{r stuck_values}
  # 30 minute filter
fb[Variance == 0 & Quality == 0, roll_sd := sd(Mean), by = list(round_minute(dateTime), variable)]
fb[roll_sd == 0 & Quality == 0 & variable == "SST"]
#not 100% sure what is going on here


ggplot(fb[roll_sd == 0 & Quality == 0]) +
  geom_point(aes(dateTime, Mean)) +
  facet_grid(variable ~ ., scales = "free_y")

```

# Validate against SmartBuoy

```{r smartbuoy_match}

sb_positions = smartbuoy.positions()[platform == "SmartBuoy" & dateTo > min(fb$dateTime)]

fb_sb = copy(fb[variable %in% c("SAL", "PAR") & Quality == 0])

for(buoy in sb_positions$deployment){
  buoypos = sb_positions[deployment == buoy,.(lon, lat)]
  fb_sb[,dist := geosphere::distGeo(data.frame(Longitude, Latitude), buoypos)]
  fb_sb[dist < 2000, deployment_group := buoy]
}
fb_sb[, dist := NULL]

sb = smartbuoy.fetch(deployment_group=unique(fb$near_sb),
                     parameters=c("SAL", "PAR"),
                     after=min(fb$dateTime), before=max(fb$dateTime))[depth < 2]

sb[, dateTime := round_minute(dateTime, 5)]
fb_sb[,.(Mean = mean(Mean)), by = list(dateTime, variable, deployment_group)]

fb_sb = merge(fb_sb, sb[,.(dateTime, deployment_group, sb_Mean = value, variable = par)],
              by = c("dateTime", "deployment_group", "variable"))

ggplot(fb_sb) +
  geom_point(aes(Mean, sb_Mean)) +
  geom_smooth(aes(Mean, sb_Mean), method = lm) +
  facet_wrap("variable", scales = "free") +
  coord_equal() + labs(x="Ferrybox", y="Buoy")

  # par outlier
fb_sb[variable == "PAR" & Mean > 1500 & sb_Mean < 750]
xlim = as.POSIXct(c("2016-08-03", "2016-08-04"))
ggplot(fb[variable == "PAR" & Quality == 0]) +
  geom_point(aes(dateTime, Mean)) +
  geom_point(data=sb[par=="PAR"], aes(dateTime, value, color = deployment_group))
```

## validate MET against ECMWF

Stream:Atmospheric model
Area:58.0N 10.0W 48.0N 8.0E
Dataset:cams_nrealtime
Step:0
Version:1
Type of level:Surface
Date:20160101 to 20161231
Grid:0.125 x 0.125
Parameter:10 metre U wind component, 10 metre V wind component, 2 metre temperature, Sea surface temperature
Class:MACC
Type:Analysis
Time:00:00:00, 06:00:00, 12:00:00, 18:00:00
Type:Forecast
Time:00:00:00, 12:00:00

```{r ECMWF}
ecmwf = read.ecmwf("ECMWF_cams_nrealtime_uk_2016.nc")
MET = copy(fb[variable %in% c("AIRT", "BAROPR", "WINDRD", "WINDRS") & Quality == 0,
              .(time = dateTime,
                longitude = Longitude,
                latitude = Latitude,
                Mean, variable, Heading)])
MET[, time := round_minute(time, 60*6)]
MET[, latitude := round(latitude/0.125)*0.125]
MET[, longitude := round(longitude/0.125)*0.125]
MET = dcast.data.table(MET, time + longitude + latitude + Heading ~ variable,
                       value.var = "Mean", fun.aggregate = mean, na.rm = T)
MET = merge(MET, ecmwf, by = c("time", "latitude", "longitude"))

ggplot(MET[month(time) == 3]) +
  geom_point(aes(time, WINDRS, color = "Endeavour")) +
  geom_point(aes(time, wsp, color = "ECMWF"))

ggplot(MET) +
  geom_point(aes(wsp, WINDRS))

MET[wsp < 1 & WINDRS > 10]
```


